\emeszida

### 2024-03-29 sigbovik/paper.tex
Added more text, one image, and some arabic script... we're getting there.

@@

\section{Introduction}
The history of math is long, but the history of programming is longer.
Cuneiform, arguably the first writing in the world, is known for its sheep receipts and beer ration lists as well as copper complaints \cite{oppenheim1954}.
This article adds neural network programming to that vaunted list of human achievements.
A newly discovered cuneiform tablet preserves the mechanism for performing simple neural network calculations.
These calculations, it seems, were used to calculate the lengths of triangles; a well known exercise preserved on the tablet known as Plimpton 322\footnote{And important and real introduction into this interesting object is found in \cite{robson2002}}.
It is remarkably that no historian of math or cuneiform scholar has ever consider this possibility.
This paper presents the story of discovery, a description of the cuneogramming language, and includes a facsimile copy of the important tablet as an appendix.

\section{Discovery}
Museum of the Sealand

Unfortunately, the hardware required to execute these programs (i.e. a living Babylonian mathematician) has not been adequately preserved, but we have managed to write a Python library which emulates it.\footnote{\url{github.com/MrLogarithm/emeszida}}

\section{Description of the Language}
Programs in \emeszida\ follow a tabular structure with three main sections: (i) a header, denoted by \textsc{dub} \textcuneiform{𒁾} (``tablet''), (ii) a sequence of instructions, and (iii) a colophon detailing the tablet's authorship.
Each instruction spans four columns, which we have taken to calling the \textit{arguments}, \textit{opcode}, \textit{destination}, and \textit{line number}. 
These columns are usually tab-separated, though in a few documents they are TAB-separated (using the cuneiform sign \textsc{tab} \textcuneiform{𒋰}).
Instructions are grouped into blocks by means of horizontal lines.

@@

@@

\section{Description of the Texts}
- add graph plotting perceptron output vs true trigonometric ratios from Plimpton

@@

\section*{Acknowledgments}

>>>>

\usepackage{soul}

\setmainfont[Ligatures=TeX]{Times New Roman}
\newfontfamily\arabicfont
    [Script=Arabic,     % to get correct arabic shaping
    Scale=1.2]          % make the arabic font bigger, a matter of taste
        {Arial}     % whatever Arabic font you like

\newcommand{\textarabic}[1]     % Arabic inside LTR
    {\bgroup\textdir TRT\arabicfont #1\egroup}
\newcommand{\n}         [1]     % for digits inside Arabic text
    {\bgroup\textdir TLT #1\egroup}
\newcommand{\afootnote} [1]     % Arabic footnotes
    {\footnote{\textarabic{#1}}}
\newenvironment{Arabic}     % Arabic paragraph
    {\textdir TRT\pardir TRT\arabicfont}{}

@@

\section{Introduction}
The history of math is long, but the history of programming is longer.
Cuneiform, arguably the first writing in the world, is known for its sheep receipts and beer ration lists as well as copper complaints \cite{oppenheim1954}.
This article adds neural network programming to that vaunted list of human achievements.
A set of newly discovered cuneiform tablets preserve the mechanism for performing simple neural network calculations.
These methods, it seems, were used to calculate the lengths of triangles; a well known exercise preserved on the tablet known as Plimpton 322\footnote{And important and real introduction into this interesting object is found in \citet{robson2002}.}.
It is remarkable that no historian of math or cuneiform scholar has ever consider this possibility.
This paper covers the background, a description of the cuneogramming language, and includes a facsimile copy of the most important tablet as an appendix.

Unfortunately, the hardware required to execute these programs (i.e. a living Babylonian mathematician) has not been adequately preserved, but we have managed to write a Python library which emulates it.\footnote{\url{github.com/MrLogarithm/emeszida}}
The assumption is that these calculations were done by hand in their copious free time between inventing the wheel and the concept of zero.
While the actual output of these tablets is relatively simple by modern standards, the implications of this discovery are profound.
Future work will explore how these techniques could have been used in the realm of astronomical calculation and elucidate the full extent of the Babylonian's computational prowess.

\section{Description of the Language}
Programs in \emeszida\ (\textsc{eme.šid.a} "language of counting") follow a tabular structure with three main sections: (i) a header, denoted by \textcuneiform{𒁾} (\textsc{dub}  "tablet"), (ii) a sequence of instructions, and (iii) a colophon detailing the tablet's authorship.
Each instruction spans four columns, which we have taken to calling the \textit{arguments}, \textit{opcode}, \textit{destination}, and \textit{line number}. 
These columns are usually tab-separated, though in a few documents they are TAB-separated (using the cuneiform sign \textsc{tab} \textcuneiform{𒋰}).
Instructions are grouped into blocks by means of horizontal lines.

@@

% Added here to end up on next page
\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{BM_34580.jpg}
  \caption{BM 34580, courtesy of the Trustees of the British Museum, CC BY-NC-SA 4.0}
  \label{fig:BM_34580}
\end{figure*}

@@

\section{Description of the Texts}
The tablets preserve a number of methods for performing simple neural network calculations.
Simple array operations as well as the calculation of the dot product are included in the corpus.
All of this comes together in the large perceptron tablet.
It is clear that the cuneogrammers were slowly developing their skills as some of the earlier and simpler methods date as far back as the Old Akkadian period (c. 2300--2150 BCE) while the perceptron tablet is from the Old Babylonian period (c. 1900--1600 BCE).
The longest tablet, included in the appendix, \hl{shows a fully featured perceptron ADD MORE HERE}
Interestingly, the long weights at the end of the tablet very closely resemble the parameters of later astronomical calculations (see Figure \ref{fig:BM_34580}).

\hl{- add graph plotting perceptron output vs true trigonometric ratios from Plimpton}

@@

\section*{Acknowledgments}
The search for these tablets and our effort to understand them was inspired by Ramsey Nasser's \textarabic{قلب} programming language (https://github.com/nasser/---).
